Parameters for UPLC data from the Lumos (used LW extracellular DOM data as test case)
Updated to use new peakShape_KL.r

Krista Longnecker
Woods Hole Oceanographic Institution
September 24, 2018
========================================================

Go ahead and load up some of the parameters:
```{r Settings, eval=TRUE}
library(xcms)
library(CAMERA)
library(RColorBrewer)
library(ggplot2)
library(ggdendro)
library(XLConnect)

```


Now go ahead and do the peak picking, or load in the existing file...peak picking can be time consuming and it easier to only do that once

```{r GettingData, eval = TRUE}

##if you want to save time and process shorter time section, can set the 
##scanrange here (min and mx time, in seconds), and add scanrange = scanrange
##to the centWave peak picking below
#scanrange=c(100,300)

usePath <- "C:/Users/krista/Documents/mzML_Lumos_MS1_updated"

ext <- ".mzML"
pre <- paste0(usePath,"/")

mzdatafiles <- list.files(usePath,recursive = FALSE, full.names=TRUE,
                          pattern = paste0("\\",ext))

# ## How many CPU cores has your machine (or cluster) ?
snowparam <- SnowParam(workers = 4, type = "SOCK")

#no need to repeat the peak picking
if(0) {
    load("ExistingDataFile.Rimage")
  } else {
    
    setNoise = 10000 #for UPLC, noise level looks ~10000 (visually, from MZmine)

    #Note on ppm - I have not converged on a good ppm for intracellular metabolites, 
    #and the extracellular value should be considered tentative for now
    xs<-xcmsSet(mzdatafiles, method = "centWave",ppm = 10,snthresh = 100,
      prefilter = c(4,setNoise), mzCenterFun = "wMean",integrate = 2, 
      verbose.columns = TRUE, peakwidth=c(5,15), fitgauss= TRUE, noise = setNoise, 
      mzdiff=-0.005,BPPARAM=snowparam) #,sleep = 0.0001)#scanrange = scanrange)
}



```


```{r settingFileNames}
# #setup the filenames all in one place bc easier...
#put this here bc if we load in data already with peak picking...that will bring in a NameOfFile

nDate <- "2018.07.10_v2"
nBase <- "dataset_UPLC_neg_samples"
fName_allPeaks <- paste0(nBase,'_listAllPeaks.',nDate,'.csv')
fName_aligned <- paste0(nBase,'_aligned.',nDate,'.csv')
fName_peaksPicked <- paste0(nBase, '_picked.',nDate, '.pdf')
fName_TICplot <- paste0(nBase,'_TICplot.',nDate,'.pdf')
NameOfFile <- paste0(nBase,"_",nDate,".Rimage")
rm(nDate,nBase)

save.image(NameOfFile)


```



```{r apply_peakShape_KL, eval=TRUE}
# # #####the peakShape code seems to be from Tony Larson...via the Google Groups forum  
# # Note the difference syntax with read_chunk from within the R Markdown file
# updated 9/24/2018 to use peakShape_KL.r which allows for the larger peak areas from the Lumos
source("peakShape_KL.r")  

if(1){
  #run peak shape...
  xs2 <- peakShape_KL(xs,cor.val=0.85)
  } else {
    #skip for the moment
    xs2<-xs
}
# 

```


Setup the ability to rename the classes based on information in the CSV file

```{r getSampleInformation, eval=TRUE}
#figure out how to pass values to/from functions later...use this to match the files 
#with the information about what sample was run for each file
wDir <-  "C:/Users/sequence_fromMethods/"

all <- readWorksheetFromFile(paste0(wDir,"EditedSequencFile.xlsx"),sheet = 1)
putDataHere <-data.frame()

for(i in 1:length(all$File.Name)) {
  h <- all$File.Name[i]
  #paste0 will strip out the spaces; make the full filename here    
  h <-paste0(pre,h,ext)
  
  #this will pick up the right row in mzdatafiles
  m <- match(basename(h),basename(xs2@filepaths))
  if (!is.na(m)) {
    tData <- all[i,]
    if (nrow(putDataHere)==0) {
      putDataHere <- tData
      } else {
    putDataHere[m,] <-tData
    }
    rm(tData)
  }   
  rm(h,m)
}   

rm(all) #keep this clean...

# rm(classlist) 
classlist <- matrix()
#this next bit is a hack, but it works in the context of xcms
for (i in 1:nrow(putDataHere)) {
  classlist[i] <- paste0(putDataHere$ionMode[i], ".", putDataHere$goodData[i])
}
rm(i)


```


```{r setupGrouping, eval=TRUE}
     
##now I will need to consider the groups. First 'group' is actually the ionization mode
#for simplicity, start by using split to get the files from one ionization mode  
#useXS <- xs2

#define the classes after the peak picking...
sampclass(xs2) <- classlist 

#note this will select files in negative ion mode that are marked with '1' in the GoodData column
xsOne <- split(xs2,classlist)[["negative.1"]]
putDataHere.1 <- split(putDataHere,classlist)[["negative.1"]]
nSamples<-length(xsOne@filepaths)

classlist.1 <- matrix()
#this next bit is a hack, but it works in the context of xcms
for (i in 1:nrow(putDataHere.1)) {
  classlist.1[i] <-paste0(putDataHere.1$ionMode[i],".",putDataHere.1$goodData[i])
}

#set the groups to the QC/blank/samples before grouping:
sampclass(xsOne) <-classlist.1

```

```{r RT_Group, eval = TRUE}

#start with the retention time correction for this set of samples

#set the center sample so it selects one of the pooled samples...which date?
#setCenter <- grep("mtab_DeepDOM_mega_FTLC_99",putDataHere.1$File.Name) #first batch

#setting the response=1 here essentially makes this a linear fit. Setting to 100 uses 100 anchors set along the RT space.
rc.obi <-retcor.obiwarp(xsOne,plottype="deviation",profStep = 0.1,distFunc = "cor", 
                         gapInit = 0.3,gapExtend=2.4) #,
                         #center = setCenter)    

xg <-group.density(rc.obi,minfrac=0.25,minsamp=1,bw=10,mzwid=0.002)     

xgF <- fillPeaks(xg, method = "chrom")     

save.image(NameOfFile)

```



```{r plotTIC, eval = FALSE}
source("getTICs_KL.r")
#xgN3 can be replaced with rc.obi
#getTICs_KL(xcmsSet = xgN3[,c(6,12,18,29)], pdfname = "TICs_finalall_pos6.pdf",rt = "corrected")

#plotting to check if peaks are aligned well
getTICs_KL(xcmsSet = xgF, pdfname = fName_TICplot, rt = "corrected")

```



This next section had some issues with the CRAM data, and I did not manage to resolve them. Also issue with full set of LW samples.

```{r makePDFofPickedPeaks, eval = FALSE}

xs.fill <- xgF #must be grouped

# 
xeic.raw <- getEIC(xs.fill, rt = "raw", groupidx= 1:nrow(xs.fill@groups))
xeic.corrected <- getEIC(xs.fill, rt = "corrected", groupidx= 1:nrow(xs.fill@groups))

#colorList should be as long as the number of possible levels in the dataset...
colorList <- c("deepskyblue","magenta","forestgreen","darkorchid","firebrick", "gold")

#plot first feature from diffreport (not sorted after pval!!)
cairo_pdf(file = fName_peaksPicked,onefile=TRUE)
for (i in 1:nrow(xs.fill@groups)){
  #for (i in 1:5){
  
  par(mfrow= c(2,1))
  #rt row data
  plot(xeic.raw, xs.fill,groupidx=i,col=colorList)
  #rt corrected data
  plot(xeic.corrected, xs.fill,groupidx=i,col=colorList)
  
}
dev.off()

```

Go ahead and spit out some variables to the folder

```{r SaveData, eval = TRUE}

#this will spit out the list of all features identified
write.csv(xgF@peaks,file = fName_allPeaks)

```

These are the pieces for CAMERA
      
```{r CAMERA, eval = TRUE}
#
#nSamples <- length(mzdatafiles)
#xsa<-xsAnnotate(xgF,sample = c(1:nSamples))
#xsa<-xsAnnotate(xgF,sample = 34) randomly try one sample for now
xsa<-xsAnnotate(xgF)

#group the features initially just by retention time
xsaF <-groupFWHM(xsa)

#figure out which features also have a matching 13C feature. Have to enter both 
#the relative error (ppm) and the absolute error (mzabs)
xsaFI <-findIsotopes(xsaF,ppm=3,mzabs = 0.01,minfrac = 1/nSamples,intval = c("into"))

#now group by the correlations based on (1) intensity, (2) EIC, (3) isotopes...
xsaC <-groupCorr(xsaFI,cor_eic_th=0.75,pval=0.05, graphMethod="hcs",
 calcIso = TRUE, calcCiS = TRUE, calcCaS = FALSE)

#setup the file to also look for adducts, only go with the primary adducts for the moment
file <-system.file("rules/primary_adducts_neg.csv",package = "CAMERA")
rules <-read.csv(file)
an <-findAdducts(xsaC,polarity = "negative",rules=rules,ppm=3)

#do some housecleaning
rm(xsa,xsaF,xsaFI,xsaC)
```


And save the final steps

```{r lastSave, eval = TRUE}
#can plot some variables       
#windows() #brings up a box for plotting that is outside RStudio (works better here)
#plotEICs(an,pspec=2,maxlabel=5)
#plotPsSpectrum(xsaC,2,maxlabel=5)

#export the data to the CSV file

#the write.csv at line 282 will only work if I ran CAMERA:
if(0) {
    #do nothing
  } else {
    write.csv(getPeaklist(an),file=fName_aligned) 
}

save.image(NameOfFile)
#


```

